---
title: "Example_Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example_Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Read package

```{r setup, message = FALSE}
library(randomForest)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rsample)
library(here)
library(purrr)
```

### Dataset

Question aim: To predict which trees are maintained by the San Francisco Department of Public Works (DPW) by using random forest.

Data: TidyTuesday 2020/01/28 San Francisco Tree. This dataset explores the difference species of trees and the legal status, site, address, etc. of them in San Francisco. The data is gathered by San Francisco Department of Legal Work.

Data dictionary: <https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-28#data-dictionary>

Data: <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-28/sf_trees.csv>

### Load Data

```{r load data}
# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}
#tuesdata <- tidytuesdayR::tt_load("2020-01-28")
#sf_trees <- tuesdata$sf_trees
if (!file.exists(here("data", "tuesdata_sf_trees.RDS"))) {
    tuesdata <- tidytuesdayR::tt_load("2020-01-28")
    sf_trees <- tuesdata$sf_trees

    # save the files to RDS objects
    saveRDS(tuesdata$sf_trees, file = here("data", "tuesdata_sf_trees.RDS"))
}
#load the data
sf_trees <- readRDS(here("data", "tuesdata_sf_trees.RDS"))
as_tibble(sf_trees)
```

### Exploratory Data Analysis

```{r parsing, warning=FALSE}
trees_df <- sf_trees %>%
  mutate(
    legal_status = case_when(
      legal_status == "DPW Maintained" ~ legal_status,
      TRUE ~ "Other"),plot_size = parse_number(plot_size)) %>%
  select(-address) %>%
  select(-species) %>%
  na.omit() %>%
  mutate_if(is.character, factor)

trees_df %>%
select_if(is.numeric) %>% 
  map(~summary(.))

map_chr(.x = trees_df, .f = typeof)

```

```{r exploratory data analysis,fig.width=7, fig.height=6}
trees_df %>%
  ggplot(aes(longitude, latitude, color = legal_status)) +
  geom_point(size = 0.3, alpha = 0.6) + 
  labs(
    title = 'Distribution of Trees by Legal Status',  
    subtitle = 'San Francisco Trees',  
    caption = "Data source: TidyTuesday 20200128",
    x = 'Longitude',  
    y = 'Latitude',  
    color = 'Legal Status'  
  ) +
  theme_minimal(base_size = 12) + 
  theme(
    legend.position = 'right',  
    plot.title = element_text(face = "bold", hjust = 0.5), 
    plot.subtitle = element_text(hjust = 0.5)  
  ) +
  scale_color_brewer(palette = "Set2")
```

```{r data analysis 2,fig.width=7, fig.height=6}
trees_df %>%
  count(legal_status, caretaker) %>%
  add_count(caretaker, wt = n, name = "caretaker_count") %>%
  filter(caretaker_count > 50) %>%
  group_by(legal_status) %>%
  mutate(percent_legal = n / sum(n)) %>%
  ggplot(aes(percent_legal, caretaker, fill = legal_status)) +
  geom_col(position = "dodge") +
  labs(
    title = "Proportion of Trees by Legal Status and Caretaker",
    subtitle = "Considering only caretakers managing over 50 trees",
    caption = "Data source: TidyTuesday 20200128",
    x = "% of Trees in Each Category",
    y = "Caretaker",
    fill = "Legal Status"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  scale_fill_brewer(palette = "Set2")

```

```{r data analysis 3,warning=FALSE,fig.width=7, fig.height=6}

sf_trees %>%
  select(tree_id, latitude, longitude) %>%
  gather(measure, value, -tree_id) %>%
  ggplot(aes(1, value)) +
  geom_boxplot(aes(fill = measure), outlier.colour = "sandybrown") +  
  scale_y_continuous(n.breaks = 30) +
  facet_wrap(~ measure, scales = "free_y") +
  labs(
    title = "Distribution of Latitude and Longitude for Trees",
    subtitle = "Comparative Analysis of Tree Positions in San Francisco",
    caption = "Data source: TidyTuesday 20200128",
    x = "Geographic Coordinate",
    y = "Value") + 
  theme_minimal(base_size = 12) +
  theme(
    strip.background = element_rect(fill = "palegreen3"),
    strip.text.x = element_text(size = 10, color = "black", face = "bold"),
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.ticks.x = element_blank(),
    panel.grid.major = element_blank(),
    #panel.grid.minor = element_blank(),
    legend.position = "none"
  )

```

## Building Prediction Models

```{r modeling,fig.width=7, fig.height=6}
set.seed(1)
trees_split <- initial_split(trees_df, strata = legal_status)
trees_train <- training(trees_split)
trees_test <- testing(trees_split)
#the best model from trying
rf_model <- randomForest(legal_status ~ ., 
                         data = trees_train, 
                         mtry = 9, 
                         ntree = 1000, 
                         nodesize = 2)

predictions <- predict(rf_model, trees_test)
confusionMatrix <- table(trees_test$legal_status, predictions)
print(confusionMatrix)
#plot the model
par(mar = c(5, 4, 4, 2) + 0.1)
plot(rf_model, main = "Random Forest Model")
mtext("on prediction of trees' legal status in SF", side = 3, line = 0.5, adj = 0.5, cex = 1)
mtext("Data source: TidyTuesday 20200128", side = 1, line = 4, adj = 1, cex = 0.8)

#Prediction testing error rate: 
print(error_rate<-(352+197)/(4328+197+352+1084))
```

Prediction testing error rate: 0.092.

## Summary

In summary, from the first plot, we can see that DPW maintained trees appeared to be in more order. They are planted by the streets of San Francisco. From the second plot, we can see that most of these trees' caretaker is private, interestingly, some of the DPW maintained trees' caretaker is not DPW. From the third plot, we can see the median latitude of the trees is around 38, the longitude is around -122.5, both with a few outlier. Our random forest model predicts which trees are maintained by the San Francisco Department of Public Works (DPW) by using random forest. After adjusting for the parameters, I found the best model for fitting, with a testing error rate of 0.092. The last plot shows our model, by using the plot function from the random forest package.

## Functions

|     | dplyr       | purrr     | ggplot2        | randomForest                    |
|------------|------------|------------|------------|------------------------|
| 1   | select()    | map()     | geom_boxplot() | randomForest()                  |
| 2   | mutate()    | map_chr() | facet_wrap()   | plot.randomForest(), aka plot() |
| 3   | na.omit()   |           | geom_point()   |                                 |
| 4   | gather()    |           | geom_col()     |                                 |
| 5   | group_by()  |           |                |                                 |
| 6   | filter()    |           |                |                                 |
| 7   | mutate_if() |           |                |                                 |
| 8   | count()     |           |                |                                 |
